---
title: "Work with data types"
format: html
---

# Instructions

This week, you worked with three new data types in R -- factors, strings, and date-times -- using the forcats, stringr, and lubridate packages, respectively.

In this milestone, you'll work with the arrival date data contained in `hotels` to determine whether there is a seasonal effect for cancelations.

To learn more about the variables contained in the data set, refer to the data dictionary: https://rstudio.github.io/academy-tidyverse-hotel-cancelations/assets/hotels_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false
```

## Data types

In this milestone, you'll parse the arrival dates, extract the month of each date, and then create a variable that records whether each booking arrived during the spring, summer, fall, or winter. You'll then inspect whether the cancelation rate varies by season.

## Recreation

### Part 1 - Import

To begin, read the data from `hotels.csv` which lives in your `data/` directory to an object named `hotels`:

```{r}
#| label: recreation-import

```

### Part 2 - Seasons

Run the code below to see a table.

```{r}
#| label: recreate-this-1
#| message: false
solution <- readr::read_csv("data/milestone05.csv")
solution
```

Your task is to use what you've learned about the lubridate package to transform `hotels` into this table. You will need to:

1. Parse the character strings in `arrival_date` to date-times
2. Save the month of each date, as a number, in a new variable named `arrival_month`
3. Recode the values of `arrival_month` to make a new variable named `season`. Recode the values such that:

    +  Months 1, 2, 3 = "Winter"
    +  Months 4, 5, 6 = "Spring"
    +  Months 7, 8, 9 = "Summer"
    +  Months 10, 11, 12 = "Fall"

(Hint: use function `case_when()` for recoding values of `arrival_month` into `season`.)

4. Add your new variables immediately after the `arrival_date` column

Work in the code chunk below. Save the result as `seasons`; we will use it in Part 3.

```{r}
#| label: recreation-seasons

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(seasons, solution, tolerance = 1e-4)
```

## Part 3 - The role of factor orders in visualizations

Now that you have a `season` variable, you can compute and visualize summaries for each season in the same manner as in previous milestones. Run the chunk below to view a summary table.

```{r}
#| label: summary
seasons_summary <-
  seasons |>
  group_by(season) |>
  summarize(
    percent_canceled = sum(is_canceled) / n() * 100
  )

seasons_summary
```

After you have run the chunk above to create `seasons_summary`, run the chunk below to visualize the data of `seasons_summary`.

```{r}
#| label: summary-plot
#| eval: false
seasons_summary |>
  ggplot(mapping = aes(x = season, y = percent_canceled)) +
  geom_col() +
  coord_cartesian(ylim = c(0, 100)) +
  labs(title = "Cancelation rate by season")
```

This is a nice plot, but the order of seasons on the x axis is arbitrary (alphabetical). Let's give the seasons a more principled order before we publish this plot.

#### Visualization

Using the File Explorer, navigate to the `images/` directory, then click on `milestone05.png` to open an image.

You should see a plot in which the seasons are now ordered chronologically from winter to fall. Your task is to recreate this plot.

Use the chunk below to recreate this plot. Hint: turn the `season` variable into a factor with a set order.

```{r}
#| label: recreation-visualize-1

```

## Extension

Using the code chunk below, investigate a research question about this data, using the skills you have learned this week and in previous weeks of Academy. Some ideas:

1. Does the season when the reservation is *requested* affect the cancelation rate?
2. Is there an interaction effect between season and lead time?
3. [any other research question of interest]


Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. If your data contains ...
  ... date-time data, use a lubridate function to correctly format your date-time variables
  ... string data, use a stringr function to manipulate your string variables
  ... categorical data, use a forcats function to modify the factor levels of your categorical variables (reorder, recode, etc.)
3. Using a table or graph, present a meaningful subset of your modified data

```{r}
#| label: extension

```
