---
title: "Import and visualize"
format: html
---

# Instructions

Welcome to your first project milestone! In this project, you will use the **tidyverse** to investigate real data related to hotel bookings.

Hotel bookings are frequently canceled, which makes it hard for hotels to plan for the future. By the end of this project, we'd like to find a way to predict whether or not a booking will be canceled. We will split our work into milestones that build on each other to iteratively explore the data.

## How to complete a milestone

Each milestone will require you to read and edit a Quarto document, like this one. **To learn how to use Quarto in Positron and RStudio, see here: https://quarto.org/docs/get-started/hello/positron.html.**

For each milestone, you will need to:

1. **Recreate a result,** like a plot, table, or model. This will require you to use the skills you have studied that week.

2. **Extend the result.** Explore a new question about the data--this will provide an opportunity to both apply the skills you learned that week and to teach yourself something new. We will provide suggestions each week, but you are encouraged to explore any aspect of the data that interests you. Alternatively, apply and extend the skills you learned that week by working with your own data. Remember, being able to teach yourself new parts of a computer language is what will make you a truly competent open source programmer---and we are here to help.

You'll share your extension with your groupmates at the weekly meeting, and show them how they could do the same.

## The `hotels` data set

The data comes from an open hotel bookings data set from Antonio, Almeida and Nunes, 2019: https://www.sciencedirect.com/science/article/pii/S2352340918315191

To learn more about the variables contained in the data set, refer to the data dictionary: https://rstudio.github.io/academy-tidyverse-hotel-cancelations/assets/hotels_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## R Basics

In this milestone, you'll become familiar with the `hotels` data set.

## Recreation

### Part 1 - Import

In the code chunk below, use a command **from the readr package** to read in the data set.

* The data set lives in the file `hotels.csv`, which is stored in the `data/` folder in your working directory.
* Save the data set as an object named `hotels`.

_Note_: Both `read_csv()` and `read.csv()` read data from a CSV file, but they differ slightly in _how_ they read in data. Make sure you use the tool from the readr package. (Not sure which is which? Check the help documentation.)

```{r}
#| label: recreation-import

```

### Part 2 - Explore

Use R to inspect the contents of `hotels`:

- Display the contents of `hotels`
- Use `nrow()` to find the number of observations in `hotels`
- Use `mean()` to find the average lead time of hotel bookings (`lead_time`)
- Use `count()` from the dplyr package to count the numbers of canceled and not canceled bookings (`is_canceled`)
- Use `summary()` to generate a summary of the data set

*Hint:* you may want to look up the help pages for each of these functions.

Write your code in the following chunk:

```{r}
#| label: recreation

```

### Part 3 - Visualize

Using the File Explorer, navigate to the `images/` directory, then click on `milestone01.png` to open an image.

You should see a plot showing the relationship between a booking's lead time and whether or not it was canceled. Your task is to recreate this plot.

Use `ggplot()` in the chunk below to re-create the plot. _Hint_: This plot uses two geoms: a geom that draws points and a geom that draws a smooth line.

```{r}
#| label: recreation-visualize

```

### What does this plot tell us?

In this plot, every booking is plotted according to its lead time (the x axis) and whether or not it was canceled (the y axis). Points appearing at y = 1 were canceled. Points appearing at y = 0 were not.

The smooth line shows the "typical" y value for each value of lead time. We could think of it as an estimated probability that a booking with that lead time will be canceled. Since the typical y value goes towards one as lead time increases, we see that bookings with a larger lead time are more likely to be canceled.

## Extension

Using the code chunk below, investigate a research question about this data, using the visualization skills you learned this week.

Some ideas (we encourage you to choose one):

1. Your recreation focused on the relationship between cancellation and lead time -- how does cancellation relate to other factors, such as the duration of the booked stay or the number of previous cancellations?
2. Do you see a similar relationship between cancellation and lead time across bookings of different deposit types?
3. Within a particular country of interest, what patterns do you see between cancellation and other variables you have explored?
4. [any other research question of interest]

Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. Filter your data using a logical test/condition
3. Graph this data subset using at least one geom

```{r}
#| label: extension

```
