---
title: "Wrangle rows and columns"
format: html
---

# Instructions

This week, you learned several more functions from the dplyr package (pronounced DEE-ply-er). These functions help you work with the rows and columns in your data set.

You'll use these skills to investigate whether travellers with children are more (or less) likely to cancel their reservation.

To learn more about the variables contained in the data set, refer to the data dictionary: https://rsacdn.link/milestones/internal/v2/hotel-cancelations/assets/hotels_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false

# Load your packages here

```

## Create and select columns

In this milestone, you'll create a new variable that notes whether a booking's guests included children. You'll then compute the cancelation rates for bookings with children and bookings without children.

## Recreation

### Part 1 - Import

Before you begin, you will need to import your data set. Use the code chunk below to read the data from the data file `hotels.csv`, which is stored in the `data/` folder in your working directory. Be sure to save the data to an object named `hotels`.

```{r}
#| label: recreation-import

```

### Part 2 - Create and summarize

Run the code below to see a table.

```{r}
#| label: recreate-this
#| message: false
solution <- readr::read_csv("data/milestone03.csv")
solution
```

Your task is to use what you've learned about the dplyr package to transform `hotels` into this table. You will need to:

1. Create a new column named `has_children`, which contains the result of a logical test
2. Group the data by `has_children`
3. Summarize the `total_bookings` and `percent_canceled` for each group

Work in the code chunk below. Save the result as `children`.

```{r}
#| label: recreation-create

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(children, solution, tolerance = 1e-4)
```

Given the results shown in your table, do you think the presence of children will be a useful way to predict whether a booking is canceled?

## Extension

Using the code chunk below, investigate a single research question about this data, using the additional data wrangling skills you learned this week. Some ideas:

1. Combine the number of weekend nights and number of weeknights to calculate the total duration of each booking. Then, explore the relationship between booking duration and other variables of interest.
2. Which bookings only fell on weekends? Do weekend-only bookings differ from those that spanned weeknights?
3. Based on the number of booking changes, divide the hotels data into groups of your choosing (e.g. low/medium/high volatility) and explore patterns based on these groups.
4. [any other research question of interest]

Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. Create at least one new variable in your data set (using `mutate()` and/or `summarize()`), or use other dplyr functions to alter your data
3. Use your updated data set to create at least one graph and/or table

```{r}
#| label: extension

```
